Analysis: Machine Learning - OblAmt - Outliers

### 2021-11-28 code/analysis_code/analysis-ML-OA-outlier.Rmd
Outlier Analysis

---
title: "Analysis: Machine Learning - OblAmt"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---

@@

It compares the  models, and then finally fits the “best” model to the test data.

@@

* Drop the unused variables
* Perform feature engineering
* Transform Incident Month into ordinal variable
* Set the random seed to 123
* Split the dataset into 70% training, 30% testing, stratified on incidentType
* 5-fold cross validation, 5 times repeated, stratified on incidentType for the CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#requested amount filter
obl_data <- analysisdata %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, ResponseDays, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, OblAmt)

@@

@@

#use the recipe specified earlier (line 133)
```
---

@@

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
BT_res <- baguette::bag_tree(cost_complexity = tune(),
                                   tree_depth = tune(),
                                   min_n = tune()) %>%
          parsnip::set_engine("rpart",
                              times = 10) %>%
          parsnip::set_mode("regression") %>%
          tune::tune_grid(preprocessor = BT_rec,
                          resamples = obl_data_bootsplit,
                          grid = tree_grid,
                          metrics = metric_set(rmse))

@@

#vip package doesn't support BT

>>>>

---
title: "Analysis: Machine Learning - OblAmt - Outliers"
output: 
  html_document:
    theme: flatly
    toc: FALSE
---

@@

It compares the  models, and then finally fits the “best” model to the test data. It removes the $4B outlier that corresponds to Hurricane Maria in Puerto Rico in 2017.

@@

* Drop the unused variables
* Perform feature engineering
* Transform Incident Month into ordinal variable
* Remove outlier
* Set the random seed to 123
* Split the dataset into 70% training, 30% testing, stratified on incidentType
* 5-fold cross validation, 5 times repeated, stratified on incidentType for the CV folds
* Create a recipe for data and fitting that codes categorical variables as dummy variables
```{r data setup}
#requested amount filter
obl_data <- analysisdata %>%
              dplyr::select(state, IncidentYear, IncidentMonth, declarationType, incidentType, IncidentDuration, ResponseDays, IH, PA, HM,
                            Population, Counties, TotalAgencies, FEMACSAvg, OblAmt)

@@

#remove Hurricane Maria outlier
obl_data <-subset(obl_data, OblAmt< 4000000000)

@@

#use the recipe specified earlier
```
---

@@

### 4. Tuning Using Cross-Validation and the `tune_grid()` function
```{r}
#tune the model with previously specified cross-validation and RMSE as target metric
BT_res <- BT_wflow %>%
          tune::tune_grid(resamples = obl_data_bootsplit,
                          grid = tree_grid,
                          metrics = metric_set(rmse),
                          control = control_resamples(save_pred = TRUE,
                                                      extract = function(x) extract_model(x)))

@@

#bag roots function to extract fit
bag_roots <-  function(x){
  x %>% 
    dplyr::select(.extracts) %>% 
    unnest(cols = c(.extracts)) %>% 
    dplyr::mutate(models = map(.extracts,
                               ~.x$model_df)) %>% 
    dplyr::select(-.extracts) %>% 
    unnest(cols = c(models)) %>% 
    mutate(root = map_chr(model,
                          ~as.character(.x$fit$frame[1, 1]))) %>%
    dplyr::select(root)  
}

#plot most important variables
bag_roots(BT_res) %>% 
  ggplot(mapping = aes(x = fct_rev(fct_infreq(root)))) + 
  geom_bar() + 
  coord_flip() + 
  labs(x = "root", y = "count")
